\chapter{Dispatching and Invoking}

\versionadded{1.1}

\ZSI{} is focused on parsing and generating SOAP messages, and provides
limited facilities for dispatching to the appropriate message handler.
This is because \ZSI{} works within many client and server environments,
and the dispatching styles for these different environments can be
very different.

Nevertheless, \ZSI{} includes some dispatch and invocation functions.
To use them, they must be explicitly imported, as shown in the example
at the start of this document.

The implementation (and names) of the these classes reflects the orientation
of using SOAP for remote procedure calls (RPC).

Both client and server share a class that defines the mechanism a
client uses to authenticate itself.

\begin{classdesc}{AUTH}{}
This class defines constants used to identify how the client
authenticated: \code{none} if no authentication was provided;
\code{httpbasic} if HTTP basic authentication was used, or
\code{zsibasic} if \ZSI{} basic authentication (see below)) was used.
\end{classdesc}

The \ZSI{} schema (see the last chapter of this manual)
defines a SOAP header element, \code{BasicAuth}, that
contains a name and password.
This is similar to the HTTP basic authentication header, except
that it can be used independently from an HTTP transport.

\section{Dispatching}

The \module{ZSI.dispatch} module allows you to expose Python functions as a
web service.
The module provides the infrastructure to parse the request, dispatch
to the appropriate handler, and then serialize any return value
back to the client.
The value returned by the function will be serialized back to the client.
To return multiple values, return a list.

If an exception occurs, a SOAP fault will be sent back to the client.

Three dispatch mechanisms are provided: one supports standard CGI
scripts, one runs a dedicated server based on the
\module{BaseHTTPServer} module, and the third uses the JonPY package,
\url{http://jonpy.sourceforge.net}, to support FastCGI.

\begin{methoddesc}{AsCGI}{\optional{module_list}}
This method parses the CGI input and invokes a function that has the
same name as the top-level SOAP request element.
The optional \code{module_list} parameter can specify a list of modules
(already imported) to search for functions.
If no modules are specified, only the \module{__main__} module will be searched.
\end{methoddesc}

\begin{methoddesc}{AsServer}{\optional{**keywords}}
This creates a \class{HTTPServer} object with a request handler that only
supports the ``POST'' method.
Dispatch is based solely on the name of the root element in the
incoming SOAP request;
the request URL is ignored.

The following keyword arguments may be used:

\begin{tableiii}{l|c|p{30em}}{textrm}{Keyword}{Default}{Description}
\lineiii{\code{docstyle}}{\code{0}}{If true, then all methods are
invoked with a single argument, the unparsed body of the SOAP message.}
\lineiii{\code{modules}}{\code{(__main__,)}}{List of modules containing
functions that can be invoked.}
\lineiii{\code{nsdict}}{\code{\{\}}}{Namespace dictionary to send in the
    SOAP \code{Envelope}}
\lineiii{\code{port}}{\code{80}}{Port to listen on.}
\end{tableiii}

\end{methoddesc}

\begin{methoddesc}{AsJonPy}{request=req\optional{, **keywords}}

This method is used within a JonPY handler to do dispatch.

The following keyword arguments may be used:

\begin{tableiii}{l|c|p{30em}}{textrm}{Keyword}{Default}{Description}
\lineiii{\code{request}}{\code{(__main__,)}}{List of modules containing
functions that can be invoked.}
\end{tableiii}

The following code shows a sample use:

\begin{verbatim}
import jon.fcgi
from ZSI import dispatch
import MyHandler

class Handler(cgi.Handler):
    def process(self, req):
        dispatch.AsJonPy(modules=(MyHandler,), request=req)

jon.fcgi.Server({jon.fcgi.FCGI_RESPONDER: Handler}).run() 
\end{verbatim}

\end{methoddesc}

\begin{methoddesc}{GetClientBinding}{}
More sophisticated scripts may want to use access the client binding object,
which encapsulates all information about the client invoking the script.
This function returns \code{None} or the binding information, an
object of type \class{ClientBinding}, described below.
\end{methoddesc}

\begin{classdesc}{ClientBinding}{...}
This object contains information about the client.
It is created internally by \ZSI{}.
\end{classdesc}

\begin{methoddesc}{GetAuth}{}
This returns a tuple containing information about the client identity.
The first element will be one of the constants from the \code{AUTH} class
described above.
For HTTP or \ZSI{} basic authentication, the next two elements will be
the name and password provided by the client.
\end{methoddesc}

\begin{methoddesc}{GetNS}{}
Returns the namespace URI that the client is using, or an empty string.
This can be useful for versioning.
\end{methoddesc}

\begin{methoddesc}{GetRequest}{}
Returns the \class{ParsedSoap} object of the incoming request.
\end{methoddesc}

The following attribute is read-only:

\begin{memberdesc}{environ}
A dictionary of the environment variables.
This is most useful when \method{AsCGI()} is used.
\end{memberdesc}

\section{The \module{client} module --- sending SOAP messages}

\ZSI{} includes a module to connect to a SOAP server over HTTP, send requests,
and parse the response.
It is built on the standard Python \module{httplib} and \module{Cookie}
modules.
It must be explicitly imported, as in
\samp{from ZSI.client import AUTH,Binding}.

\begin{classdesc}{Binding}{\optional{**keywords}}
This class encapsulates a connection to a server, known as a \emph{binding}.
A single binding may be used for multiple RPC calls.
Between calls, modifiers may be used to change the URL being posted to,
etc.

Cookies are also supported; if a response comes back with a \code{Set-Cookie}
header, it will be parsed and used in subsequent interactions.

The following keyword arguments may be used:

\begin{tableiii}{l|c|p{20em}}{textrm}{Keyword}{Default}{Description}
\lineiii{\code{auth}}{\code{(AUTH.none,)}}{A tuple with authentication
    information; the first value should be one of the constants
    from the \class{AUTH} class.}
\lineiii{\code{nsdict}}{\code{\{\}}}{Namespace dictionary to send in the
    SOAP \code{Envelope}}
\lineiii{\code{soapaction}}{\code{''}}{Value for the
    \code{SOAPAction} HTTP header.}
\lineiii{\code{readerclass}}{\code{None}}{Class used to create DOM-creating
XML readers; see the description in the \class{ParsedSoap} class.}
\lineiii{\code{writerclass}}{\code{None}}{ElementProxy Class used to create 
XML writers; see the description in the \class{SoapWriter} class.}
\lineiii{\code{tracefile}}{\code{None}}{An object with a \code{write}
    method, where packet traces will be recorded.}
\lineiii{\code{transport}}{HTTPConnection/HTTPSConnection}{transport class}
\lineiii{\code{transdict}}{\{\}}{keyword arguments for connection initialization}
\lineiii{\code{url}}{n/a}{URL to post to.}
\end{tableiii}

If using SSL, the \code{cert_file} and \code{key_file} keyword parameters may
also be used.
For details see the documentation for the \module{httplib} module.

\end{classdesc}

Once a \class{Binding} object has been created, the following modifiers are
available.
All of them return the binding object, so that multiple modifiers can
be chained together.

\begin{methoddesc}{AddHeader}{header, value}
Output the specified \code{header} and \code{value} with the HTTP
headers.
\end{methoddesc}

\begin{methoddesc}{SetAuth}{style, name, password}
The \code{style} should be one of the constants from the \code{AUTH}
class described above.
The remaining parameters will vary depending on the \code{style}.
Currently only basic authentication data of name and password are
supported.
\end{methoddesc}

\begin{methoddesc}{SetNS}{uri}
Set the default namespace for the request to the specified \code{uri}.
\end{methoddesc}

\begin{methoddesc}{SetURL}{url}
Set the URL where the post is made to \code{url}.
\end{methoddesc}

\begin{methoddesc}{ResetHeaders}{}
Remove any headers that were added by \method{AddHeader()}.
\end{methoddesc}

The following attribute may also be modified:

\begin{memberdesc}{trace}
If this attribute is not \code{None}, it should be an object with a
\code{write} method, where packet traces will be recorded.
\end{memberdesc}

Once the necessary parameters have been specified (at a minimum, the URL
must have been given in the constructor are through \code{SetURL}),
invocations can be made.

\begin{methoddesc}{RPC}{url, opname, pyobj, replytype=None\optional{, **keywords}}
This is the highest-level invocation method.
It calls \method{Send()} to send \code{pyobj} to the specified \code{url}
to perform the \code{opname} operation,
and calls \method{Receive()} expecting to get a reply of the specified
\code{replytype}.

This method will raise a \exception{TypeError} if the response does not
appear to be a SOAP message, or if is valid SOAP but contains a fault.
\end{methoddesc}

\begin{methoddesc}{Send}{url, opname, pyboj\optional{, **keywords}}
This sends the specified \code{pyobj} to the specified \code{url}, invoking
the \code{opname} method.
The \code{url} can be \code{None} if it was specified in the \class{Binding}
constructor or if \code{SetURL} has been called.
See below for a shortcut version of this method.

The following keyword arguments may be used:

\begin{tableiii}{l|c|p{20em}}{textrm}{Keyword}{Default}{Description}
\lineiii{\code{auth\_header}}{\code{None}}{String (containing presumably
    serialized XML) to output as an authentication header.}
    SOAP \code{Envelope}
\lineiii{\code{nsdict}}{\code{\{\}}}{Namespace dictionary to send in the
    SOAP \code{Envelope}}
\lineiii{\code{requestclass}}{n/a}{Python \code{class} object with a
    \code{typecode} attribute specifying how to serialize the data.}
\lineiii{\code{requesttypecode}}{n/a}{Typecode specifying how to serialize
    the data.}
\lineiii{\code{soapaction}}{Obtained from the \class{Binding}}{Value for the
    \code{SOAPAction} HTTP header.}
\end{tableiii}

\end{methoddesc}

Methods are available to determine the type of response that came back:

\begin{methoddesc}{IsSOAP}{}
Returns true if the message appears to be a SOAP message.
(Some servers return an HTML page under certain error conditions.)
\end{methoddesc}

\begin{methoddesc}{IsAFault}{}
Returns true if the message is a SOAP fault.
\end{methoddesc}

Having determined the type of the message (or, more likely, assuming
it was good and catching an exception if not), the following methods
are available to actually parse the data.
They will continue to return the same value until
another message is sent.

\begin{methoddesc}{ReceiveRaw}{}
Returns the unparsed message body.
\end{methoddesc}

\begin{methoddesc}{ReceiveSoap}{}
Returns a \class{ParsedSOAP} object containing the parsed message.
Raises a \exception{TypeError} if the message wasn't SOAP.
\end{methoddesc}

\begin{methoddesc}{ReceiveFault}{}
Returns a \class{Fault} object containing the SOAP fault message.
Raises a \exception{TypeError} if the message did not contain a fault.
\end{methoddesc}

\begin{methoddesc}{Receive}{replytype=None}
Parses a SOAP message.
The \code{replytype} specifies how to parse the data.
If it s \code{None}, dynamic parsing will be used, usually resulting
in a Python list.
If \code{replytype} is a Python class, then the class's \code{typecode}
attribute will be used, otherwise \code{replytype} is taken to be
the typecode to use for parsing the data.
\end{methoddesc}

Once a reply has been parsed (or its type examined), the following
read-only attributes are available.
Their values will remain unchanged until another reply is parsed.

\begin{memberdesc}{reply_code}
The HTTP reply code, a number.
\end{memberdesc}

\begin{memberdesc}{reply_headers}
The HTTP headers, as a \class{mimetools} object.
\end{memberdesc}

\begin{memberdesc}{reply_msg}
A text string containing the HTTP reply text.
\end{memberdesc}

Finally, if an attribute is fetched other than one of those described
above, it is taken to be the \code{opname} of a remote procedure,
and a callable object is returned.
This object dynamically parses its arguments, receives the reply, and
parses that.

\begin{methoddesc}{opname}{args...}
Using this shortcut requires that the \method{SetURL()} was invoked first.
This method is then equivalent to:
\samp{RPC(None, opname, tuple(args), TC.Any())}
\end{methoddesc}

