% $Header$
\documentclass{manual}
\title{ZSI: The Zolera Soap Infrastructure}
\author{Rich Salz}
\authoraddress{
DataPower Technology, \url{http://www.datapower.com}\\
E-mail: \email{rsalz@datapower.com}\\
}
\date{March 12, 2003}
\input{version}

\begin{document}

\maketitle

\centerline{\strong{ABSTRACT}}

\module{ZSI}, the Zolera SOAP Infrastructure, is a Python package that
provides an implementation of SOAP messaging, as described in
\citetitle[http://www.w3.org/TR/soap]{The SOAP 1.1 Specification}.
In particular, \module{ZSI} parses and generates SOAP messages, and
converts between native Python datatypes and SOAP syntax.
It can also be used to build applications using
\citetitle[http://www.w3.org/TR/SOAP-attachments]{SOAP Messages with
Attachments}.
\module{ZSI} is ``transport neutral'', and provides only a simple
I/O and dispatch framework; a more complete solution is the
responsibility of the application using \module{ZSI}.
As usage patterns emerge, and common application frameworks are
more understood, this may change.

\module{ZSI} requires Python 2.0 or later and PyXML version 0.6.6 or later.

The \module{ZSI} homepage is at
\url{http://pywebsvcs.sf.net/}.

\vfill

\centerline{\strong{COPYRIGHT}}

Copyright \copyright{} 2001, Zolera Systems, Inc.\\
All Rights Reserved.

Permission is hereby granted, free of charge, to any person obtaining a
copy of this software and associated documentation files (the "Software"),
to deal in the Software without restriction, including without limitation
the rights to use, copy, modify, merge, publish, distribute, and/or
sell copies of the Software, and to permit persons to whom the Software
is furnished to do so, provided that the above copyright notice(s) and
this permission notice appear in all copies of the Software and that
both the above copyright notice(s) and this permission notice appear in
supporting documentation.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT
OF THIRD PARTY RIGHTS. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR HOLDERS
INCLUDED IN THIS NOTICE BE LIABLE FOR ANY CLAIM, OR ANY SPECIAL INDIRECT
OR CONSEQUENTIAL DAMAGES, OR ANY DAMAGES WHATSOEVER RESULTING FROM LOSS
OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT, NEGLIGENCE
OR OTHER TORTIOUS ACTION, ARISING OUT OF OR IN CONNECTION WITH THE USE
OR PERFORMANCE OF THIS SOFTWARE.

Except as contained in this notice, the name of a copyright holder
shall not be used in advertising or otherwise to promote the sale, use
or other dealings in this Software without prior written authorization
of the copyright holder.

\vfill

\centerline{\strong{Acknowledgments}}

We are grateful to the members of the \code{soapbuilders}
mailing list (see \url{http://groups.yahoo.com/soapbuilders}),
Frederick Lundh for his \code{soaplib} package (see
\url{http://www.secretlabs.com/downloads/index.htm\#soap}),
and Cayce Ullman and Brian Matthews for their \code{SOAP.py} package
(see \url{http://sourceforge.net/projects/pywebsvcs}).

\tableofcontents{}
\def\ZSI{\module{ZSI}}

\input{c01-intro}
\input{c02-samples}
\input{c03-except}
\input{c04-utils}
\input{c05-parse}
\input{c06-tc}
\input{c07-writer}
\input{c08-fault}
\input{c09-resolve}
\input{c10-dispatch}
\input{c11-wsdl}

\chapter{ZSI Schema}

The \ZSI{} schema defines two sets of elements.  One is used to enhance
the SOAP Fault \code{detail} element, and to report header errors.
The other is used to define a header element containing a name and
password, for a class of basic authentication.

% tell latex2html to load the verbatim package so we get \verbatiminput
\ifx\htmlonly\undefined\else\usepackage{verbatim}\fi
\verbatiminput{zsixsd.tex}

\end{document}
